---
title: "Homework 1"
author: "Nicholas Esposito"
date: "9/15/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### Load in the data for Q1

```{r}
spdata <- read.csv("http://www.stat.ufl.edu/~winner/data/bioequiv_sulf.csv")
attach(spdata); #names(spdata)

library(knitr)
```


### P.1.A - See Written PDF for Work By-Hand

95% CI: (-177.61, 1177.09)

We can be 95% confident the mean difference of the Test formulation and the Reference formulation falls between -177.61 and 1177.09.

Since the test statistic, t = 1.4869, is less than the critical value, t(0.025,44) = 2.01537, we fail to reject the null hypothesis that µT - µR = 0.

```{r}
## Create a variable that is AUC for the drug sulfadoxine and a variable for formulation

AUC.sulf <- y[measure==2 & drug==1]       # measure=AUC, drug=sulfadoxine
form21 <- form[measure==2 & drug==1]               # form=1 if Test, form=2 if Ref
form.AUC.sulf <- factor(form21, levels=1:2, labels=c("T","R"))
cbind(form.AUC.sulf, AUC.sulf)
plot(AUC.sulf ~ form.AUC.sulf, main="Sulfadoxine AUC by Formulation")

## Compute n, ybar, sd for Test and Reference

n.T <- length(AUC.sulf[form.AUC.sulf=="T"])
n.R <- length(AUC.sulf[form.AUC.sulf=="R"])
ybarl.T <- mean(AUC.sulf[form.AUC.sulf=="T"])
ybarl.R <- mean(AUC.sulf[form.AUC.sulf=="R"])
sdl.T <- sd(AUC.sulf[form.AUC.sulf=="T"])
sdl.R <- sd(AUC.sulf[form.AUC.sulf=="R"])

## Print results in tabular form
T.stats <- cbind(n.T, ybarl.T, sdl.T)
R.stats <- cbind(n.R, ybarl.R, sdl.R)
TR.stats <- rbind(T.stats, R.stats)
colnames(TR.stats) <- c("n", "mean", "SD")
rownames(TR.stats) <- c("Test", "Reference")
round(TR.stats, 5)

## Compute pooled SD, 95%CI for muT-muR and print results

s_p <- sqrt(((n.T-1)*sdl.T^2+(n.R-1)*sdl.R^2)/(n.T+n.R-2))   # Pooled SD
df <- n.T+n.R-2                                              # Degrees of Freedom

ybarl.diff <- ybarl.T - ybarl.R                              # Mean Difference
se.diff <- s_p * sqrt(1/n.T + 1/n.R)                         # Standard Error of mean diff
t.025 <- qt(.975, df)                                          # Critical t-value for 95% CI

mul.LB <- ybarl.diff - t.025 * se.diff                        # Lower Confidence Limit
mul.UB <- ybarl.diff + t.025 * se.diff                        # Upper Confidence Limit

## Print out summary of 95%CI for muT-muR
ci.out <- cbind(df, s_p, ybarl.diff, se.diff, t.025, mul.LB, mul.UB)
colnames(ci.out) <- c("df","pooled SD", "mean diff", "Std Err", "t(.975)", "Lower Bound", "Upper Bound")
round(ci.out, 5)
```
```{r}
## Use t.test function for 95%CI for µT-µR

AUC.ttest <- t.test(AUC.sulf ~ form.AUC.sulf, var.equal=TRUE, conf.level=0.95)
AUC.ttest
```
### P.1.B - See Written PDF for Work By-Hand

95% CI: (0.4279, 2.3788)

We can be 95% confident the ratio σT^2 / σ_R^2 falls between 0.4279 and 2.3788.

Since the test statistic, F = 1.0089, is less than the critical value, F(0.975,22,22) = 2.358, we fail to reject the null hypothesis that σ_T^2 = σ_R^2.

```{r}
## Use var.test function for 95%CI for σ1^2/σ2^2

AUC.vartest <- var.test(AUC.sulf ~ form.AUC.sulf, var.equal=TRUE, conf.level=0.95)
AUC.vartest
```
### P.1.C - See Written PDF for Work By-Hand

β0 is the mean of Y (AUC value) where subjects receive the Reference formulation (X = 0).

β1 is the difference between the means (µT - µR).

µT can be found by adding β0 and β1.

When conducting the t-test for the H0: β1 = 0, we find the same test statistic (|t| = 1.4869) and the same p-value (0.1442) as the t-test in P.1.A.

```{r}
# Create var X to represent function for Test/Ref
X <- c(rep(1,23), rep(0,23))

lm_X <- lm(AUC.sulf ~ X)
summary(lm_X)

plot(AUC.sulf ~ X, main="AUC.sulf vs Test/Reference Formulation")
abline(lm_X)

X.ttest <- t.test(AUC.sulf ~ X, var.equal=TRUE, conf.level=0.95)
X.ttest
```

 
## Question 2 - See Written PDF for Work By-Hand

### All the work for Question 2 can be found in the written PDF.

## Question 3

### Read in the data for Q3

```{r}
# Read in data
spain_latlong <- read.csv(
  "https://users.stat.ufl.edu/~winner/data/spain_latlong.csv")
attach(spain_latlong); #names(spain_latlong)
head(spain_latlong)
tail(spain_latlong)
```


### P.3.A - Obtain scatterplots of WGS ref - GIS - (Y) versus Ptolemy (X) for lat and long

```{r}
# Define X and Y for lat and long
Y_long_wgs <- Long_wgs
X_long_ptol <- Long_ptol
Y_lat_wgs <- Lat_wgs
X_lat_ptol <- Lat_ptol

# Plots for Y versus X (lat and long)
par(mfrow=c(1,2))

plot(Y_long_wgs ~ X_long_ptol, main="WGS Longitude vs Ptolemy Longitude")

plot(Y_lat_wgs ~ X_lat_ptol, main="WGS Latitude vs Ptolemy Latitude")
```

### P.3.B - Fit simple linear regression models, relating GIS (Y) to Ptolemy (X)

```{r}
# Fit linear models
lm_long <- lm(Y_long_wgs ~ X_long_ptol)
lm_lat <- lm(Y_lat_wgs ~ X_lat_ptol)

# Plot scatterplots again, now with linear models fitted
par(mfrow=c(1,2))

plot(Y_long_wgs ~ X_long_ptol, main="WGS Longitude vs Ptolemy Longitude")
abline(lm_long)

plot(Y_lat_wgs ~ X_lat_ptol, main="WGS Latitude vs Ptolemy Latitude")
abline(lm_lat)
```

### P.3.C - Test whether there is a positive association between GIS and Ptolemy

#### WGS Longitude (GIS) vs Ptolemy Longitude

H0: β1 <= 0, Ha: β1 > 0 

Since we have a p-value that is very small (<2e-16), we reject the null hypothesis. Therefore, there is a positive association between WGS Longitude (GIS) and Ptolemy Longitude.

```{r}
summary(lm_long)
```

#### WGS Latitude (GIS) vs Ptolemy Latitude

H0: β1 <= 0, Ha: β1 > 0 

Since we have a p-value that is very small (<2e-16), we reject the null hypothesis. Therefore, there is a positive association between WGS Latitude (GIS) and Ptolemy Latitude.

```{r}
summary(lm_lat)
```

### P.3.D - C95% for β1 > 0

#### Linear model for WGS Longitude (GIS) vs Ptolemy Longitude

The 95% confidence interval for β1 is (0.7213,0.8684). 

We are 95% confident the true value for β1 (the mean change) falls between 0.7213 and 0.8684. 

Since the interval does not contain 0, we can say the mean change in GIS as Ptolemy 
is “increased by 1 unit”, or that the mean change is positive.

```{r}
confint(lm_long)
```

#### Linear model for WGS Latitude (GIS) vs Ptolemy Latitude

The 95% confidence interval for β1 is (0.7830,0.9498). 

We are 95% confident the true value for β1 (the mean change) falls between 0.7830 and 0.9498. 

Since the interval does not contain 0, we can say the mean change in GIS as Ptolemy 
is “increased by 1 unit”, or that the mean change is positive.

```{r}
confint(lm_lat)
```

### P.3.E ANOVAs, F-tests, coefficients of correlation and determination

#### Linear model for WGS Longitude (GIS) vs Ptolemy Longitude

Coefficient of correlation: 0.9546814

Coefficient of determination: 0.9114167

H0: σ1^2 / σ2^2 = 1, Ha: σ1^2 / σ2^2 != 1

Since the p-value obtained by the t-test (0.2128) exceeds the alpha value of 0.05, we fail to reject the null hypothesis. The true ratio of variances is not equal to 1.

```{r}
# Linear model for longitude
anova(lm_long)
aov(Y_long_wgs ~ X_long_ptol)
cor(X_long_ptol, Y_long_wgs) # X, Y

# R-squared (coeff. of determination)
summary(lm_long)$r.squared

# F-test 
long.var.test <- var.test(X_long_ptol, Y_long_wgs, var.equal=TRUE, conf.level=0.95) # X, Y
long.var.test
```

#### Linear model for WGS Latitude (GIS) vs Ptolemy Latitude

Coefficient of correlation: 0.9511858

Coefficient of determination: 0.9047545

H0: σ1^2 / σ2^2 = 1, Ha: σ1^2 / σ2^2 != 1

Since the p-value obtained by the t-test (0.5245) exceeds the alpha value of 0.05, we fail to reject the null hypothesis. The true ratio of variances is not equal to 1.

```{r}
# Linear model for latitude
anova(lm_lat)
aov(Y_lat_wgs ~ X_lat_ptol)
cor(X_lat_ptol, Y_lat_wgs) # X, Y

# R-squared (coeff. of determination)
summary(lm_lat)$r.squared
 
# F-test
lat.var.test <- var.test(X_lat_ptol, Y_lat_wgs, var.equal=TRUE, conf.level=0.95) # X, Y
lat.var.test
```








