---
title: "Bollywood Movie Regression"
author: "Larry Winner"
date: "9/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Bollywood Box Office Regression Analysis
# Read in the Dataset

```{r readdata}
bolly1 <- read.csv(
    "http://www.stat.ufl.edu/~winner/data/bollywood_boxoffice2.csv")
attach(bolly1); names(bolly1)
head(bolly1)
tail(bolly1)
```

# Consider Plots of Gross (Y) versus Budget (X) in Original and Log Scales and plot

```{r plotscale}
Y1 <- Gross; Y2 <- log(Gross)
X1 <- Budget; X2 <- log(Budget)

# win.graph(height=5.5, width=7.0)
par(mfrow=c(2,2))
plot(Y1 ~ X1, main="Gross vs Budget")
abline(lm(Y1 ~ X1))
plot(Y2 ~ X1, main="ln(Gross) vs Budget")
abline(lm(Y2 ~ X1))
plot(Y1 ~ X2, main="Gross vs ln(Budget)")
abline(lm(Y1 ~ X2))
plot(Y2 ~ X2, main="ln(Gross) vs ln(Budget)")
abline(lm(Y2 ~ X2))
par(mfrow=c(1,1))
```

# Fitting OLS model based on (natural) logs of Y and X

```{r olsfit}
Y <- Y2; X <- X2
n <- length(Y)
Xbar <- mean(X)
Ybar <- mean(Y)
SS_XX <- sum((X-Xbar)^2)
SS_XY <- sum((X-Xbar)*(Y-Ybar))
SS_YY <- sum((Y-Ybar)^2)
b1 <- SS_XY / SS_XX
b0 <- Ybar - b1 * Xbar
# Print out results
ols.out1 <- cbind(n, Xbar, Ybar, SS_XX, SS_XY, SS_YY, b0, b1)
colnames(ols.out1) <- c("n", "Xbar", "Ybar", "SS_XX", "SS_XY", "SS_YY", "b0", "b1")
round(ols.out1, 3)
```

# Compute the Predicted values, Residuals, Error Sum of Squares, and Estimated Error Variance

```{r pred_s2}
Yhat <- b0 + b1 * X
e <- Y - Yhat
SSE1 <- SS_YY - b1^2 * SS_XX
SSE <- sum(e^2)
s2 <- SSE / (n-2)
# Print Out results
ols.out2 <- cbind(b0, b1, SSE, n-2, s2)
colnames(ols.out2) <- c("b0", "b1", "SSE", "df", "s2=MSE")
round(ols.out2,3)
```

# Obtain Estimated Standard Errors for b0, b1, t-tests and 95% CI's for beta0, beta1

```{r se_t_ci}
SE_b1 <- sqrt(s2/SS_XX)
SE_b0 <- sqrt(s2*(1/n + Xbar^2/SS_XX))
SE_yhat <- sqrt(s2*(1/n + (X-Xbar)^2/SS_XX))
SE_e <- sqrt(s2*(1 - 1/n - (X-Xbar)^2/SS_XX))
b <- rbind(b0,b1)
SE_b <- rbind(SE_b0,SE_b1)
t_b <- b / SE_b
p_b <- 2*(1-pt(abs(t_b),n-2))
beta_LB <- b + qt(.025,n-2) * SE_b
beta_UB <- b + qt(.975,n-2) * SE_b
beta.out <- cbind(b, SE_b, t_b, p_b, beta_LB, beta_UB)
colnames(beta.out) <- c("Estimate", "Std. Err.", "t", "P", "LB","UB")
rownames(beta.out) <- c("Intercept", "Budget")
round(beta.out, 4)
```

# Compute 95% Confidence Interval (Mean) and Prediction Interval (Individual Movie) with Budget=100

```{r cipi}
Xstar <- log(100)
Ystar <- b0 + b1 * Xstar
SE_mean_Xs <- sqrt(s2*(1/n + (Xstar-Xbar)^2/SS_XX))
SE_pred_Xs <- sqrt(s2*(1 + 1/n + (Xstar-Xbar)^2/SS_XX))
CI_LB <- Ystar + qt(.025,n-2) * SE_mean_Xs
CI_UB <- Ystar + qt(.975,n-2) * SE_mean_Xs
PI_LB <- Ystar + qt(.025,n-2) * SE_pred_Xs
PI_UB <- Ystar + qt(.975,n-2) * SE_pred_Xs

Xstar1 <- exp(Xstar); Ystar1 <- exp(Ystar)
CI_LB1 <- exp(CI_LB); CI_UB1 <- exp(CI_UB)
PI_LB1 <- exp(PI_LB); PI_UB1 <- exp(PI_UB)

ci.pi.out <- cbind(Xstar1,Ystar1,CI_LB1,CI_UB1,PI_LB1,PI_UB1)
colnames(ci.pi.out) <- cbind("X*","Y*","CI_LB","CI_UB","PI_LB","PI_UB")
round(ci.pi.out,2)
```

# Computations Using lm function

```{r reglm}
bolly.mod1 <- lm(Y ~ X)
summary(bolly.mod1)
confint(bolly.mod1)
(ci.mean <- predict(bolly.mod1, list(X=log(100)), int="c"))    # Confidence Interval
(pi.indiv <- predict(bolly.mod1, list(X=log(100)), int="p"))   #
exp(ci.mean); exp(pi.indiv)
```








