---
title: "Take_Home_Exam_1"
author: "Nicholas Esposito"
date: "9/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in the data (x and y values)

```{r}
X <- c(130, 130, 140, 140, 150, 150, 160, 160, 170, 170, 180, 180)
Y <- c(19.68, 22.29, 19.62, 15.97, 19.17, 21.34, 18.22, 19.26, 14.68, 14.19, 13.31, 12.57)
```

1. Obtain the least squares fitted equation. Give the first fitted value and residual.

```{r}
lm_1 <- lm(Y ~ X)
lm_1

fitted(lm_1)[1]
residuals(lm_1)[1]
```

2. Give the error sum of squares and the mean square error.

```{r}
n <- length(X) # Number of observations

lm_1_SSE <- sum((fitted(lm_1)-Y)^2)
lm_1_MSE <- lm_1_SSE / (n-2)

lm_1_SSE
lm_1_MSE
```

3. Give a 95% confidence interval for b1

```{r}
confint(lm_1)
```

4. Give a 95% confidence interval for b0

```{r}
confint(lm_1)
```
5. Give the 95% confidence interval for E{Y|X=150}

```{r}
X_value <- data.frame(X=c(150))
predict(lm_1, newdata = X_value, interval="confidence", level=0.95)

ssXX <- sum((X-155)^2)
sY <- sqrt(lm_1_MSE*(1/12 + 25/ssXX))

yhat <- 18.26529
upr <- yhat + (sY * 2.228)
lwr <- yhat - (sY * 2.228)

upr
lwr
```

6. Give the 95% prediction interval for a new observation when X=150

```{r}
predict(lm_1, X_value, interval="prediction", level=0.95)

ssXX <- sum((X-155)^2)
sY <- sqrt(lm_1_MSE*(1 + 1/12 + 25/ssXX))

yhat <- 18.26529
upr <- yhat + (sY * 2.228)
lwr <- yhat - (sY * 2.228)

upr
lwr
```

7. Give the Analysis of Variance

```{r}
anova(lm_1)

#F(0.95) = 4.9646
```

8. Use the F-test to test H0: b1 = 0 versus Ha: b1 != 0 (alpha = 0.05)

```{r}

```

9. Obtain R^2 and r

```{r}
summary(lm_1)

cor(X,Y)
```

10. Use the correlation test (see notes) to test H0: p = 0 versus Ha: p != 0 (alpha = 0.05)

```{r}
cor.test(X,Y) # Pearson's correlation test
```




